---
title: "Get HUC-12 Codes and Elevations for All Point Sources"
author: Lindsay Poirier
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

The purpose of the code presented in this document is to compose the final datasets that will be used to analyze carceral facility proximity to potential PFAS sites. Towards this end, the code:

1. Loads source datasets and transforms them to a common CRS (4269)
2. Determines which watershed boundary (HUC-12) each facility in the source datasets geo-spatially intersect with
3. Determines the elevation of each facility in the source datasets via the USGS Elevation Point Query Service
4. Gets the census block for each carceral facility and classifies the census block as urban or rural
5. Saves all new datasets as .gpkg files for use in `carceral_analysis.Rmd`, `hospitals_analysis.Rmd`, `nursing_homes_analysis.Rmd`, and `schools_analysis.Rmd`

> Notes: New files are saved often throughout the code to prevent having to re-execute computationally heavy areas should RStudio get hung up. All files that are unnecessary for `carceral_analysis.Rmd` are unlinked eventually. 

## Load Packages

```{r}
library(tidyverse)
library(sf)
library(rgeos)
library(readxl)
library(elevatr)
library(tigris)
```

## Load Source Datasets

### Census Block Urban Areas

```{r}
ua <- read.delim("https://www2.census.gov/geo/docs/reference/ua/2020_UA_BLOCKS.txt",
                 header = TRUE, 
                 sep = "|") %>%
  select(GEOID) %>%
  mutate(URBAN = 1)
```


### Carceral Facilities

> Source: Department of Homeland Security. "Prison Boundaries." Homeland Infrastructure Foundation Level Data, July 7, 2020. [https://hifld-geoplatform.opendata.arcgis.com/datasets/geoplatform::prison-boundaries/about](https://hifld-geoplatform.opendata.arcgis.com/datasets/geoplatform::prison-boundaries/about)

```{r}
pb_sf <- st_read("source_datasets/Prison_Boundaries-shp/Prison_Boundaries.shp") %>%
  st_transform(crs = 4269) %>%
  st_transform(crs = 32617) %>% #convert to utm for calculating centroids
  st_centroid() %>% #centroids from original multipolygons
  st_transform(crs = 4269) %>% 
  mutate(INDEX_CARCERAL = row_number())
```
### Public Schools

> Downloaded June 27, 2023

```{r}
schools_sf <- st_read("source_datasets/Public_Schools/PublicSchools.shp") %>%
  st_transform(crs = 4269) %>%
  st_transform(crs = 32617) %>% #convert to utm for calculating centroids
  st_centroid() %>% #centroids from original multipolygons
  st_transform(crs = 4269) %>% #back to 4269
  mutate(Index = row_number())
```

### Nursing Homes

> Downloaded June 27, 2023

```{r}
nursing_homes_sf <- st_read("source_datasets/Nursing_Homes/Nursing_Homes.shp") %>%
  st_transform(crs = 4269) %>%
  st_transform(crs = 32617) %>% #convert to utm for calculating centroids
  st_centroid() %>% #centroids from original multipolygons
  st_transform(crs = 4269) %>% #back to 4269
  mutate(Index = row_number())
```
### Hospitals

> Downloaded June 27, 2023

```{r}
hospital_sf <- st_read("source_datasets/Hospitals/Hospitals.shp") %>%
  st_transform(crs = 4269) %>%
  st_transform(crs = 32617) %>% #convert to utm for calculating centroids
  st_centroid() %>% #centroids from original multipolygons
  st_transform(crs = 4269) %>% #back to 4269
  mutate(Index = row_number()) 
```

### Presumptive PFAS Contamination

> Salvatore D, Mok K, Garrett KK, et al. Presumptive Contamination: A New Approach to PFAS Contamination Based on Likely Sources. Environ Sci Technol Lett. 2022;9(11):983-990. doi:10.1021/acs.estlett.2c00502

```{r}
susp_cont <- read_csv("source_datasets/suspected_contamination_220712.csv") %>%
  select(-1) %>%
  rename(Index = X) %>%
  st_as_sf(coords=c("Longitude","Latitude"), crs = 4326) %>%
  st_transform(crs = 4269) %>%
  mutate(STATE = state.abb[match(State,state.name)]) %>%
  mutate(REGION = case_when(
    STATE %in% c("CA", "WA", "OR", "HI", "AK") ~ "PACIFIC",
    STATE %in% c("AZ", "CO", "ID", "NM", "MT", "UT", "NV", "WY") ~ "MOUNTAIN",
    STATE %in% c("IA", "KS", "MN", "MO", "NE","ND", "SD") ~ "WEST NORTH CENTRAL",
    STATE %in% c("IN", "IL", "MI", "OH", "WI") ~ "EAST NORTH CENTRAL",
    STATE %in% c("AR", "LA", "OK", "TX") ~ "WEST SOUTH CENTRAL",
    STATE %in% c("AL", "KY", "MS", "TN") ~ "EAST SOUTH CENTRAL",
    STATE %in% c("NJ", "NY", "PA") ~ "MID ATLANTIC",
    STATE %in% c("MA", "CT", "RI", "VT", "NH", "ME") ~ "NEW ENGLAND",
    STATE %in% c("DE", "DC", "GA", "MD", "NC", "SC", "VA", "WV", "FL") ~ "SOUTH ATLANTIC",
    TRUE ~ "OUTLYING"
    ))
```

### Known PFAS Contamination

> PFAS Project Lab. PFAS Contamination Site Tracker. Published June 2022. Accessed August 11, 2022. https://pfasproject.com/pfas-sites-and-community-resources/

```{r}
pfas_sf <- read.csv("source_datasets/pfas.csv")  %>%
  filter(!is.na(latitude) | !is.na(longitude)) %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326, na.fail = FALSE) %>%
  st_transform(4269)
```

### TRI

> Downloaded August 24, 2023
> This is the preliminary 2022 TRI data file, last updated August 16, 2023. The EPA estimates that this file is 99% complete. However, during the months of August and September, they will update the dataset based on late submissions and revisions to the data. 

```{r}
tri_sf <- read.csv("source_datasets/tri_2022_us.csv") |>
  filter(X45..PFAS == "YES") |>
  select(X3..FRS.ID, X4..FACILITY.NAME, X12..LATITUDE, X13..LONGITUDE) |>
  distinct() %>%
  st_as_sf(coords = c("X13..LONGITUDE", "X12..LATITUDE"), crs = 4269, na.fail = FALSE)
```


## Summaries of facility totals

### How many facilities were analyzed in each category?

This code block produces a table listing the number of facilities that were analyzed in each category.

```{r}
susp_cont %>%
  st_drop_geometry() %>%
  group_by(Source) %>%
  summarize(Count = n())
```

### How many carceral facilities were not open or missing population data?

The Department of Homeland Security extracts information about facility operating statuses, capacities, and current populations from city, county, and state websites and documents. That data is missing from this dataset if it hasn't been recorded via these sources. 

```{r}
pb_sf %>%
  st_set_geometry(NULL) %>%
  summarize(num_not_closed = sum(STATUS != "CLOSED"),
            percent_not_closed = sum(STATUS != "CLOSED") / n() * 100,
            num_missing_pop = sum(POPULATION == -999),
            percent_missing_pop = sum(POPULATION == -999) / n() * 100)
  
```

## Get HUC-12s Boundary Codes and Elevations for Each Point Source

> The HUC-12 Watershed Boundary Dataset is not included in the GitHub repo due to size (102992 rows). It can be accessed at [https://water.usgs.gov/GIS/huc.html](https://water.usgs.gov/GIS/huc.html)

```{r}
#wbd <- st_read("source_datasets/WBD_National_GDB/WBD_National_GDB.gdb/")
#check_validity <- st_is_valid(wbd)

#When we ran this analysis, ten points from the watershed boundary file had invalid geometry. We used functions from the sf package to make the geometries valid. 

#wbd_valid <- st_make_valid(wbd)
#rm(wbd)

#st_write(wbd_valid, "source_datasets/wbd_valid.gpkg")
wbd_valid <- st_read("source_datasets/wbd_valid.gpkg")
```


For each point source, we first determined where point locations intersected with watershed boundaries. We added the corresponding HUC-12 code to the point source data. Then, using the elevatr package (which streamlines access to the USGS Elevation Point Query Service API), we determined the elevations for each facility. 

### Get HUC and Elevation for Carceral Facilities

```{r}
intersected_pb_wbd <- st_intersects(pb_sf, wbd_valid)

huc_pb_sf <- pb_sf %>%
  mutate(intersection = as.integer(intersected_pb_wbd),
         HUC = if_else(is.na(intersection), "",
                        wbd_valid$huc12[intersection]))

elevation_pb <- get_elev_point(huc_pb_sf, prj = 4269, src = "epqs")

elevation_pb_lat_long <- 
  elevation_pb %>% 
  mutate(LONGITUDE = unlist(map(elevation_pb$geom,1)),
         LATITUDE = unlist(map(elevation_pb$geom,2)))

get_census <- function(row_num) {

  print(paste("Getting census info for row", row_num))
  elevation_pb_lat_long %>%
    slice(row_num) %>% 
    mutate(GEOID = call_geolocator_latlon(LATITUDE, LONGITUDE))
}

block_pb <-  
  map_dfr(1:nrow(elevation_pb_lat_long), get_census)

block_pb <-
  block_pb %>%
  mutate(GEOID = as.numeric(GEOID)) %>%
  left_join(ua, by = "GEOID") %>%
  mutate(URBAN = case_when(URBAN == 1 ~ 1, 
                           TRUE ~ 0))
#st_write(elevation_pb, "huc_elevation_datasets/pb_HUC_elevation.gpkg")
st_write(block_pb, "huc_elevation_datasets/pb_HUC_elevation_block.gpkg")

#rm(pb_sf, intersected_pb_wbd, huc_pb_sf)
```

### Get HUC and Elevation for Schools

```{r}
intersected_schools_wbd <- st_intersects(schools_sf, wbd_valid)

huc_schools_sf <- schools_sf %>%
  mutate(intersection = as.integer(intersected_schools_wbd),
         HUC = if_else(is.na(intersection), "",
                        wbd_valid$huc12[intersection]))

#Elevation API requests are handled in chunks to avoid timeouts.

get_elev <- function(data){
  
  row_num <- data %>% 
    st_drop_geometry() %>%
    slice(1) %>% 
    select(Index) %>% 
    pull()
  
  print(paste("Collecting elevation for data starting at row", row_num))
  
  elevation_schools_x <- 
    get_elev_point(data, 
                   prj = 4269, 
                   src = "epqs")
  #st_write(elevation_schools_x, paste0("huc_elevation_datasets/schools_HUC_elevation", row_num, ".gpkg"))
}

n <-  5000 

huc_schools_sliced <- 
  huc_schools_sf %>% 
  group_by(row_number() %/% n) %>% 
  group_map(~ .x)

elevation_schools <- 
  map_dfr(huc_schools_sliced, 
          get_elev)

st_write(elevation_schools, "huc_elevation_datasets/schools_HUC_elevation.gpkg")
```

### Get HUC and Elevation for Nursing Homes

```{r}
intersected_nursing_homes_wbd <- st_intersects(nursing_homes_sf, wbd_valid)

huc_nursing_homes_sf <- nursing_homes_sf %>%
  mutate(intersection = as.integer(intersected_nursing_homes_wbd),
         HUC = if_else(is.na(intersection), "",
                        wbd_valid$huc12[intersection]))

get_elev <- function(data){
  
  row_num <- data %>% 
    st_drop_geometry() %>%
    slice(1) %>% 
    select(Index) %>% 
    pull()
  
  print(paste("Collecting elevation for data starting at row", row_num))
  
  elevation_nursing_homes_x <- 
    get_elev_point(data, 
                   prj = 4269, 
                   src = "epqs")
  st_write(elevation_nursing_homes_x, paste0("huc_elevation_datasets/nursing_homes_HUC_elevation", row_num, ".gpkg"))
}

n <-  3000 

huc_nursing_homes_sliced <- 
  huc_nursing_homes_sf %>% 
  group_by(row_number() %/% n) %>% 
  group_map(~ .x)

elevation_nursing_homes <- 
  map_dfr(huc_nursing_homes_sliced, 
          get_elev)

st_write(elevation_nursing_homes, "huc_elevation_datasets/nursing_homes_HUC_elevation.gpkg")
rm(nursing_homes_sf, intersected_nursing_homes_wbd, huc_nursing_homes_sf)
```

### Get HUC and Elevation for Hospitals

```{r}
intersected_hospital_wbd <- st_intersects(hospital_sf, wbd_valid)

huc_hospital_sf <- hospital_sf %>%
  mutate(intersection = as.integer(intersected_hospital_wbd),
         HUC = if_else(is.na(intersection), "",
                        wbd_valid$huc12[intersection]))

get_elev <- function(data){
  
  row_num <- data %>% 
    st_drop_geometry() %>%
    slice(1) %>% 
    select(Index) %>% 
    pull()
  
  print(paste("Collecting elevation for data starting at row", row_num))
  
  elevation_hospital_x <- 
    get_elev_point(data, 
                   prj = 4269, 
                   src = "epqs")
  #st_write(elevation_hopital_x, paste0("huc_elevation_datasets/nursing_homes_HUC_elevation", row_num, ".gpkg"))
}

n <-  3000 

huc_hospital_sliced <- 
  huc_hospital_sf %>% 
  group_by(row_number() %/% n) %>% 
  group_map(~ .x)

elevation_hospital <- 
  map_dfr(huc_hospital_sliced, 
          get_elev)

elevation_hospital_lat_long <- 
  elevation_hospital %>% 
  mutate(LONGITUDE = unlist(map(elevation_hospital$geom,1)),
         LATITUDE = unlist(map(elevation_hospital$geom,2)))

get_census <- function(row_num) {

  print(paste("Getting census info for row", row_num))
  elevation_hospital_lat_long %>%
    slice(row_num) %>% 
    mutate(GEOID = call_geolocator_latlon(LATITUDE, LONGITUDE))
}

block_hospital <-  
  map_dfr(1:nrow(elevation_hospital_lat_long), get_census)

block_hospital <-
  block_hospital %>%
  mutate(GEOID = as.numeric(GEOID)) %>%
  left_join(ua, by = "GEOID") %>%
  mutate(URBAN = case_when(URBAN == 1 ~ "URBAN", 
                           TRUE ~ "RURAL")) 

#st_write(elevation_hospital, "huc_elevation_datasets/hospital_HUC_elevation.gpkg")
st_write(block_hospital, "huc_elevation_datasets/hospital_HUC_elevation_block.gpkg")
rm(hospital_sf, intersected_hospital_wbd, huc_hospital_sf)
```

### Get HUC and Elevation for Presumptive PFAS Sites

```{r}
intersected_susp_cont_wbd <- 
  st_intersects(susp_cont, wbd_valid)

huc_susp_cont_sf <- 
  susp_cont %>%
  mutate(intersection = as.integer(intersected_susp_cont_wbd),
         HUC = if_else(is.na(intersection), "",
                        wbd_valid$huc12[intersection]))

#Elevation API requests are handled in chunks to avoid timeouts.

get_elev <- function(data){
  
  row_num <- data %>% 
    st_drop_geometry() %>%
    slice(1) %>% 
    select(Index) %>% 
    pull()
  
  print(paste("Collecting elevation for data starting at row", row_num))
  
  elevation_susp_cont_x <- 
    get_elev_point(data, 
                   prj = 4269, 
                   src = "epqs")
}

n <-  5000 

huc_susp_cont_sf_sliced <- 
  huc_susp_cont_sf %>% 
  group_by(row_number() %/% n) %>% 
  group_map(~ .x)

elevation_susp_cont <- 
  map_dfr(huc_susp_cont_sf_sliced, 
          get_elev)

st_write(elevation_susp_cont, "huc_elevation_datasets/susp_cont_HUC_elevation.gpkg")

rm(susp_cont, intersected_susp_cont_wbd, huc_susp_cont_sf_sliced, huc_susp_cont_sf)
```

### Get HUC and Elevation for Known PFAS Sites

```{r}
intersected_pfas_wbd <- st_intersects(pfas_sf, wbd_valid)

huc_pfas_sf <- pfas_sf %>%
  mutate(intersection = as.integer(intersected_pfas_wbd),
         HUC = if_else(is.na(intersection), "",
                        wbd_valid$huc12[intersection]))

elevation_pfas <- get_elev_point(huc_pfas_sf, prj = 4269, src = "epqs")

st_write(elevation_pfas, "huc_elevation_datasets/pfas_HUC_elevation.gpkg")
rm(pfas_sf, intersected_pfas_wbd, huc_pfas_sf)
```

### Get HUC and Elevation for TRI Sites

```{r}
intersected_tri_wbd <- st_intersects(tri_sf, wbd_valid)

huc_tri_sf <- tri_sf %>%
  mutate(intersection = as.integer(intersected_tri_wbd),
         HUC = if_else(is.na(intersection), "",
                        wbd_valid$huc12[intersection]))

elevation_tri <- get_elev_point(huc_tri_sf, prj = 4269, src = "epqs")

st_write(elevation_tri, "huc_elevation_datasets/tri_HUC_elevation.gpkg")
rm(tri_sf, intersected_tri_wbd, huc_tri_sf)
```
